---
title: Comparing intervention effects on zero-calorie beverages consumption
author: Yian Lin, Haibo Wu, Wakeel Adekunle Kasali
date: 2024-03-01
format: 
  pdf :
    fontsize: 11pt
number-sections: true
header-includes: 
  \usepackage{caption}
  \newcommand{\beginAppendix}{
    \setcounter{table}{0}  
    \renewcommand{\thetable}{A\arabic{table}}}
bibliography: ../references/hw2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)
#devtools::install_github("stat545ubc-2023/rmiss")
library(rmiss)
library(knitr)
library(here)
```

## Introduction
Sugary beverage consumption poses health concerns because of its association with several health challenges like obesity and diabetes. Different intervention methods are proposed to steer consumers toward zero-calorie beverages that are thought healthier. This study aims to investigate the efficacy of five different intervention methods, three of which focus on visual calorie messaging via posters, and the other two focus on price discount. To assess and compare different intervention methods, sales of zero-calorie beverages and sugary beverages under different interventions were monitored in three different sites. 

The statistical questions of interest are: 

1. Do the interventions significantly influence the sales of zero-calorie and sugary beverages?
2. Which interventions are more effective? Specifically, does the combination of interventions lead to larger effects? Does presenting calorie-equivalent messaging have larger impacts than presenting the simple calorie admonishment?
3. Does the effectiveness of different interventions vary between sites?

The structure of the remainder of this report is as follows. Following a brief description about the dataset in Section 2, suggestions on exploratory data analysis (EDA) are presented in Section 3. Section 4 proposes formal analysis methods, and Section 5 summarizes our recommendations. 

## Data Summary
This is a quasi-experiment study for investigating the effects of five interventions  on consumer inclination towards zero-calorie beverages. The dataset contains the daily sales of different types of beverages (including zero-calorie, sugared and others) under various interventions across three sites, from October 27 to May 23. In this study, the daily sales of each type of beverage are used for measuring the consumers' preference on this type of beverage. The crucial variables are the daily sales of zero-calorie beverages, sugared beverages and total of all the beverages, which are measured as continuous data. On the other hand, based on the objectives of the study, the site and intervention should be treated as categorical variables for further analysis, and the days of week are also recorded as a categorical variable to determine potential impacts of time on the outcomes. Additionally, the dataset includes a time-counting variable that captures the timeline of the entire experiment, indicating that each site has around 220 time points. For the entire dataset, the records of these key variables are relatively complete, with sugared and zero-calorie beverages having nine observations missing. However, records for other types of beverages show a certain proportion of missing data, which could be because they are not the items of interest and excluded in recording.

## EDA Analysis
Before doing any analysis, it is important to first assess data missingness. A graph (like Figure 1) can be drawn to visualize the proportion of missing values for all columns in the dataset. If there are missing data, it is suggested to create a table to show the data missingness by `Site` for columns that have missing data (see Table 1). To see for each site from which intervention type the missing data comes from, an expanded table can be made to show the data missingness by `Site` and `Intervention` for columns with missing data (see Table A1 in Appendix). For columns that will be used in the analysis to answer the research question, it is important to find out why the data are missing.
\
```{r}
sale_data <- read.table(here("rawdata/june1data.csv"), head=T, sep=",")

# replace "dis " with "dis"
sale_data <- sale_data %>%
  mutate(Intervention = replace(Intervention, Intervention == "dis ", "dis"))

#convert DofW, Site and Intervention columns to factor
sale_data <- sale_data %>% 
  mutate_at(c('DofW', 'Site', 'Intervention'), as.factor)
```

```{r missing-data-plot, fig.height=3}
#missing data, overall
plot_missing_values_proportions(df=sale_data) +
  ggtitle("", subtitle = "Figure 1: The proportion of missing values for all columns in \nthe dataset.")
```

```{r missing-data-table}
# data missingness  by Site
count_missing<- sale_data %>%
    group_by(Site) %>%
    summarise(across(.cols = where(~ any(is.na(.x))), ~sum(is.na(.x))))

# Table
kable(count_missing, caption = "Number of missing observations for columns with missing data present by site")
```

To explore the change of daily sales of zero-calorie beverages and bottled sugared beverages over time, time series plot for `ZeroCal` and `Sugary` can be made for each of the three sites. It is important to note that the increase in the raw counts of say zero-calorie beverages being sold does not necessarily mean the decrease in the sales of sugared beverages. It could be the case that more customs came to the shops that day so that sales of all things tended to be higher. Therefore, in addition to visualize the change of daily sales using the raw counts, it is even more important to explore the change of daily sales in percentage terms (i.e. `ZeroCal`/(`ZeroCal`+`Sugary`)\*100\% and `Sugary`/(`ZeroCal`+`Sugary`)\*100\%)
(see Figure 2 as an example). Change in the zero-calorie beverages percentage can better represent the change in zero-calorie beverages consumption relative to the sugared beverages consumption, and vice versa. 

```{r}
# delete rows with missing data in both ZeroCal and Sugary
sale_data_c <- sale_data %>%
  filter(!(is.na(ZeroCal) | is.na(Sugary)))
```

To compare the consumption of a selected type of beverages (zero-calorie beverages or sugary beverages) under different `intervention` categories, boxplots with sales in percentage terms as y-axis and with `intervention` as x-axis can be made (see Figure 3 as an example). Similarly, boxplots can also be used to compare the consumption of a selected type of beverages among days of week to see if there are any patterns.
\
```{r time-series-plot, fig.height=3.8, fig.width=7}
sale_data_c$Sugary_percent <-
  sale_data_c$Sugary/(sale_data_c$Sugary+sale_data_c$ZeroCal)*100
sale_data_c$ZeroCal_percent <-
  sale_data_c$ZeroCal/(sale_data_c$Sugary+sale_data_c$ZeroCal)*100

# Plot Daily sales of ZeroCal_percent
ggplot(data=sale_data_c, 
       aes(x=Count, y=ZeroCal_percent, group=Site)) +
  geom_line(aes(color=Intervention)) + 
  geom_point(aes(color=Intervention), size=0.5) + 
  facet_wrap( ~ Site, ncol = 1, scales = "free_y") + 
  xlab("Day") + 
  ylab("Daily percentage of \nzero-calorie beverages sales (%)") +
  ggtitle("",subtitle = "Figure 2: Daily percentage of zero-calorie beverages sales over time for \neach of the three sites")
```


```{r boxplot}
#Create a new Intervention column to make boxplots
sale_data_c$Intervention_new <- sale_data_c$Intervention
# replace the "preint" in the 2nd half of the study with "preint2"
# replace the "follow" in the 2nd half of the study with "follow2"
levels(sale_data_c$Intervention_new) <- c(levels(sale_data_c$Intervention_new), "preint2", "follow2") #first increase the levels of Intervention_new

sale_data_c <- sale_data_c %>%
  mutate(Intervention_new = replace(Intervention_new, Count>=110 & Intervention_new == "preint", "preint2"), 
         Intervention_new = replace(Intervention_new, Count>=110 & Intervention_new == "follow", "follow2"))

# reorder the levels
sale_data_c$Intervention_new <- fct_relevel(sale_data_c$Intervention_new,
                                            c("preint", "dismes", "dis", "follow", "preint2", "cal", "excer", "both", "follow2", "wash", "wash2"))

#boxplots
ggplot(data=sale_data_c %>% 
         filter(Intervention_new!="wash" & Intervention_new !="wash2"), 
       aes(x=Intervention_new, y=ZeroCal_percent)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, ncol = 3, scales = "free_y") + 
  xlab("Intervention") + 
  ylab("Daily percentage of \nzero-calorie beverages sales (%)") +
  ggtitle("",subtitle = "Figure 3: Boxplots of daily percentage of zero-calorie beverages sales \nfor each of the three sites") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


## Formal Analysis
To see which interventions lead to increases in zero-calorie beverages consumption and/or decreases in sugary beverages consumption, it is recommended to use generalized linear mixed models (GLMMs), which introduce random effects to the corresponding generalized linear model and are applicable to count longitudinal data.The individual-specific random effects in GLMMs can incorporate the between-individual variations and capture the correlation in the repeated outcomes over a period of time. The major assumption of GLMMs is also the normality of random variables. 

It is recommended to fit GLMM by introducing the random effects to sites, because the considerable differences can be seen (Figure 2) in the tendency of beverage consumption among the three sites, and the random effects can account the variabilities between different sites. The count of zero-calorie beverages or sugary beverages are response variables. This count data directly aligns with the objectives to quantify changes in consumer behavior by assessing the actual volume of beverage sales under different conditions. 

Regarding the predictor variables, we recommend using `Intervention` and `DofW`. For `Intervention`, it is important to first rename some of the categories. For the observations made in the second pre-intervention period and the second follow-up period, the values of `Intervention` should be respectively replaced by something like "preint2" and "follow2". This way we are not mixing up the first and second pre-intervention and follow-up period in our analysis. Also, "wash" and "wash2" can be combined into one category as they are not of the main interest in the study. For `DofW`, Monday through Friday can be merged into a category called weekday, and Saturday and Sunday can be merged into a category called weekend. When adding text label of `DofW` to each of the data points in Figure 2, it can be seen that most of the valleys of the time series happen at weekends. Therefore, it is important to control for day of week when making comparisons between intervention and pre-intervention periods. In summary, the non-residual part of each site-specific model can be specified as ZeroCal_count ~ Intervention_updated + DofW_updated + site or Sugary_percent ~ Intervention_updated + DofW_updated +site.

Another crucial point needs further consideration in GLMM is the choice of link function. For modeling count data, the log link function is commonly used with the Poisson or negative binomial distribution [@fox]. The log link ensures that the predicted counts are always positive.The choice can be theoretically justified based on the distribution of the response variable and the nature of the variables being modeled. Moreover, in this case, model fit statistics (e.g., AIC) are suggested to determine the most suitable link function. For example, the GLMM with the smallest AIC values can be chosen as the final model. Based on the results of the final model, the estimated coefficients of different interventions can be used for intervention effect assessment and comparison. On the other hand, the estimated random effects of each site can illustrate the affects caused by location variability.

One limitation of this approach is the challenge in statistically identifying the differences in intervention effects across different sites. However, we could still learn differences in random effects between sites in a sense that maybe statistically significant  are detected for some interventions in some sites but not in other sites. 

## Conclusions
It is recommended to use daily sales in percentage terms as response variable for the analysis, to better represent the relative consumption of zero-calorie and sugared beverages. Besides the tables showing the missingness of the dataset, the time series plots can be displayed to visualize the daily changes of zero-calorie drinks consumption, and the side by side boxplots can illustrate the relationship between the daily sales and various interventions at each site. 

We suggest to implement GLS for each site to capture the association between changes in beverage consumption and the adopted interventions, allowing us to account for the dependence structure of residuals of the model.  Specifically, `Intervention` and `DofW` should serve as the predictor variables. The optimal model for each site can be determined by comparing the AIC across several candidate models, with each model specifying one of four different correlation structures. Differences in the effectiveness of various interventions and sites are identifiable by comparing the coefficients of the final models. 

\newpage
## References {.unnumbered}
::: {#refs}
:::

\newpage
## Statistical Appendix {.unnumbered}
\beginAppendix

In this appendix, we mainly focus on implementing the suggested models to approach the proposed statistical questions, and also provide detailed interpretations for the results. 

### Model Construction
To develop a suitable GLMM for this dataset, the primary consideration should be the selection of random effects. The dataset clearly shows significant variability in beverage consumption across the three sites, indicating that incorporating a random intercept for each site in the model is a reasonable approach, which allows the model to capture the variation in beverage consumption attributable to differences among the sites. Another consideration is setting random slopes for intervention within site to allow the effect of interventions to vary across sites, e.g., some interventions might be more effective in certain sites than in others, reflecting the unique contexts or characteristics of those sites. Moreover, the selection of the link function also requires careful consideration. Both the Poisson and the negative binomial link functions are commonly utilized in (GLMMs) when dealing with count data as the response variable. Due to the negative binomial model's additional dispersion parameter, the estimates can be more reliable when the data exhibit overdispersion. However, determining which method aligns more closely with the data's nature requires further analysis. The AIC can be employed to select the most suitable model for the final analysis, providing a quantitative measure to compare the relative quality of different models. Generally, we suggest to construct models with the following formation,


 

```{r}
# data missingness  by Site
count_missing<- sale_data %>%
    group_by(Site, Intervention) %>%
    summarise(across(.cols = where(~ any(is.na(.x))), ~sum(is.na(.x))))

# Table
kable(count_missing, caption = "Number of missing observations for columns with missing data present by site and intervention type.")
 
```

